---
output:
  pdf_document: default
  html_document: default
---

# Super Learning {#sl3-advanced}

Based on the [`sl3` `R` package](https://github.com/tlverse/sl3) by _Jeremy
Coyle, Nima Hejazi, Ivana Malenica, Rachael Phillips, and Oleg Sofrygin_.

Updated: `r Sys.Date()`

## Learning Objectives{-}
1. Estimate the cross-validated predictive performance of the ensemble super
   learner.
2. Evaluate `sl3` variable importance metrics.
3. Flexibly estimate conditional density functionals with `sl3`.
4. TODO

## Cross-validated super learner

We can cross-validate the SL to see how well the SL performs on unseen data, and
obtain an estimate of the cross-validated risk of the SL.

This estimation procedure requires an outer/external layer of
cross-validation, also called nested cross-validation, which involves setting
aside a separate holdout sample that we don’t use to fit the SL. This external
cross-validation procedure may also incorporate 10 folds, which is the default
in `sl3`. However, we will incorporate 2 outer/external folds of
cross-validation for computational efficiency.

We also need to specify a loss function to evaluate SL. Documentation for the
available loss functions can be found in the [`sl3` Loss Function
Reference](https://tlverse.org/sl3/reference/loss_functions.html).

```{r CVsl}
washb_task_new <- make_sl3_Task(
  data = washb_data,
  covariates = covars,
  outcome = outcome,
  folds = origami::make_folds(washb_data, fold_fun = folds_vfold, V = 2)
)
CVsl <- CV_lrnr_sl(
  lrnr_sl = sl_fit, task = washb_task_new, loss_fun = loss_squared_error
)
CVsl %>%
  kable(digits = 4) %>%
  kableExtra:::kable_styling(fixed_thead = T) %>%
  scroll_box(width = "100%", height = "300px")
```
<!-- Explain summary!!!! -->

## Variable importance measures

Variable importance can be interesting and informative. It can also be
contradictory and confusing. Nevertheless, we like it, and so do our
collaborators, so we created a variable importance function in `sl3`! The `sl3`
`importance` function returns a table with variables listed in decreasing order
of importance (i.e., most important on the first row).

The measure of importance in `sl3` is based on a risk ratio, or risk difference,
between the learner fit with a removed, or permuted, covariate and the learner
fit with the true covariate, across all covariates. In this manner, the larger
the risk difference, the more important the variable is in the prediction.

The intuition of this measure is that it calculates the risk (in terms of the
average loss in predictive accuracy) of losing one covariate, while keeping
everything else fixed, and compares it to the risk if the covariate was not
lost. If this risk ratio is one, or risk difference is zero, then losing that
covariate had no impact, and is thus not important by this measure. We do this
across all of the covariates. As stated above, we can remove the covariate and
refit the SL without it, or we just permute the covariate (faster)
and hope for the shuffling to distort any meaningful information that was
present in the covariate. This idea of permuting instead of removing saves a lot
of time, and is also incorporated in the `randomForest` variable importance
measures. However, the permutation approach is risky, so the importance function
default is to remove and refit.

Let's explore the `sl3` variable importance measurements for the `washb` data.

```{r varimp, eval = FALSE}
washb_varimp <- importance(sl_fit, loss = loss_squared_error, type = "permute")
washb_varimp %>%
  kable(digits = 4) %>%
  kableExtra:::kable_styling(fixed_thead = TRUE) %>%
  scroll_box(width = "100%", height = "300px")
```

```{r varimp-plot, out.width = "100%", eval = FALSE}
# plot variable importance
importance_plot(
  washb_varimp,
  main = "sl3 Variable Importance for WASH Benefits Example Data"
)
```
<!-- Explain summary!!!! -->

## Categorical outcome prediction

## Multivariate outcome prediction

## Super learning of a conditional density
<!--
OPTIONAL NOTES TO CONSIDER ON THIS TOPIC FROM MARK'S VIDEO LECTURE:
Suppose we want to construct a Super Learner of the conditional probability
distribution $g_0(a\mid W)=P_0(A=a\mid W)$, where $a\in {\cal A}$.
Let's denote the values of $a$ with $\{0,1,\ldots,K\}$. A valid loss function
for the conditional density is
\[
L(g)(O)=-\log g(A\mid W).\]
That is, $g_0=\arg\min_g P_0L(g)$, i.e., $g_0$ is the minimizer of the
expectation of the log-likelihood loss.

**Candidate estimators**

1. Candidate estimators based on multinomial logistic regression: To start
with, one can use existing parametric model based MLE and machine learning
algorithms in `R` that fit a multinomial regression. For example, parametric
model multinomial logistic regression is available in `R` so that one can
already build a rich library of such estimators based on  different candidate
parametric models. In addition, `polyclass()` is a multinomial logistic
regression machine learning algorithm in `R`.

2. Candidate estimators based on machine learning for multinomial logistic
regression: Secondly, one can use a machine learning algorithm such as
`polyclass()` in `R` that data adaptively fits a multinomial logistic
regression, which itself has tuning parameters, again generating a class of
candidate estimators.

3. Incorporating screening: Note that one can also marry any of these choices
with a screening algorithm, thereby creating more candidate estimators of
interest. The screening can be particularly important when there are many
variables.

4. Candidate estimators by fitting separate logistic regressions and using
post-normalization

* Code $A$ in terms of Bernoullis $B_k=I(A=k)$, $k=0,\ldots,K$.
* Construct an estimator $\bar{g}_{nk}$ of $\bar{g}_{0k}(W)\equiv P_0(B_k=1\mid
  W)$ using any of the logistic regression algorithms, for all $k=0,\ldots,K$.
* This implies an estimator
\[
g_n(a\mid W)=\frac{\bar{g}_{na}(W)}{\sum_{k=0}^K \bar{g}_{nk}(W)}.\]
* In other words, we simply normalize these separate logistic regression
estimators so that we obtain a valid conditional distribution.
* This generates an enormous amount of interesting algorithms, since we have
available the whole machine learning literature for binary outcome regression.

5. Candidate estimators by estimating the conditional "hazard" with pooled
logistic regression.
Note that
\[
g_0(a\mid W)=\lambda_0(a\mid W) S_0(a\mid W),\]
where \[
\lambda_0(a\mid W)=P_0(A=a\mid A\geq a,W),\]

and $S_0(a\mid W)=\prod_{s\leq a}(1-\lambda_0(s\mid W))$ is the conditional
survival function $P_0(A>a\mid W)$. So we have now parameterized the
conditional distribution of $A$, given $W$, by a conditional hazard
$\lambda_0(a\mid W)$: $g_0=g_{\lambda_0}$.

* We could now focus on constructing candidate estimators of
$\lambda_0(a\mid W)$, which implies candidate estimators of $g_0$.

* For every observation $A_i$, we can create $A_i+1$ rows of data
$(W,s,I(A_i=s))$, $s=0,\ldots,A_i$, $i=1,\ldots,n$. We now run a logistic
regression estimator based on the pooled data set, ignoring ID, where we
regress the binary outcome $I(A_i=s)$ on the covariates $(W,s)$.

* If one assumes a parametric model, then this is nothing else then using the
maximum likelihood estimator, demonstrating that ignoring the ID is not
inefficient.

* This defines now an estimator of $\lambda_0(s\mid W)=P_0(A=s\mid W,A\geq s)$
as a function of $(s,W)$.

* Different choices of logistic regression based estimators will define
different estimators.

* The pooling across $s$ is not very sensible if $A$ is not an ordered variable
If $A$ is categorical, we recommend to compute  a separate logistic regression
estimator of $\lambda_0(a\mid W)$ for each $a$ (i.e., stratify by $s$ in the
  above pooled data set).

* For non-categorical $A$, one could include both stratified (by level) as well
as pooled (across levels) based logistic regression estimators.
-->

In many contexts it may be useful to estimate the conditional density of a given
random variable, given certain variables that precede them in time. In the
context of causal inference, this arises most readily when working with
continuous-valued treatments, in which case the classical propensity score
$\mathbb{P}(A = 1 \mid W)$ becomes instead the conditional density of treatment
$A$, given covariates $W$, often called the _generalized propensity score_. The
estimation of such quantities often requires specialized approaches tied to
very specific algorithmic implementations. To our knowledge, general and
flexible algorithms for conditional density estimation (CDE) have been proposed
only sparsely in the literature -- we are aware of and have implemented two such
approaches in `sl3`. The first of these is a semiparametric conditional density
estimation approach that makes certain assumptions about the constancy of
(higher) moments of the underlying distribution, while the second exploits the
relationship between the (conditional) hazard and density functions to allow CDE
via pooled hazard regression. Both approaches are flexible in that they allow
the use of arbitrary regression functions or machine learning algorithms for the
estimation of nuisance quantities (the conditional mean or the conditional
hazard, respectively). We elaborate on these two frameworks below. Importantly,
when considering the use of CV to select an optimal CDE approach or to create an
ensemble of several candidate CDE methods, we stress that, per
@dudoit2005asymptotics and related works, a loss function appropriate for
density estimation is the negative log-density loss $L(\cdot) = -\log(p_n)$.

## Moment-restricted Location-scale Conditional Density Estimation

This family of semiparametric CDE approaches exploits the general form $\rho(Y -
\mu(X) / \sigma(X))$, where $\rho$ is a given marginal density function and
$\mu(X)$ and $\sigma(X)$ are nuisance quantities that may be estimated flexibly.
CDE procedures formulated within this framework may be characterized as
belonging to a _conditional location-scale_ family, that is, in which
$p_n(Y \mid X) = \rho((Y - \mu_n(X)) / \sigma_n(X))$. Here, the marginal
density mapping $\rho$ is selected _a priori_, leaving only the relevant moments
$\mu(X)$ and $\sigma(X)$ to be estimated. While the limitation to (conditional)
location-scale families is not without its potential disadvantages (leading to
possible misspecification bias), this strategy is flexible in that it allows for
arbitrary machine learning algorithms to be used in estimating $\mu(X) = \E(Y
\mid X)$ and, optionally, the conditional variance $\sigma(X) = \E[(Y -
\mu(X))^2 \mid X]$.

In settings with limited data, the additional structure imposed by the
assumption that the target density belongs to a location-scale family may prove
advantageous, though, in practice, it is impossible to know whether and when
this assumption holds. This procedure is not a novel contribution of our own
(and we have been unable to locate a formal description of it in the
literature); nevertheless, we provide an informal algorithm sketch below. This
algorithm considers access to $n$ copies of an observed data random variable
$O = (Y, X)$, an _a priori_-specified kernel function specification $\rho$, a
candidate regression procedure $f_{\mu}$ to estimation the conditional mean
$\mu(X)$, and a candidate regression procedure $f_{\sigma}$ to estimate the
conditional variance $\sigma(W)$.

1. Estimate $\mu(X) = \E[Y \mid X]$, the conditional mean of $Y$ given $X$, by
   applying the regression estimator $f_{\mu}$, yielding $\hat{\mu}(X)$.
2. Estimate $\sigma(X) = \mathbb{V}[A \mid W]$, the conditional variance of $Y$
   given $X$, by applying the regression estimator $f_{\sigma}$, yielding
   $\hat{\sigma}^2(X)$. Note that this step involves only estimation of the
   conditional mean $\E[(Y - \hat{\mu}(X))^2 \mid X]$.
3. Estimate the one-dimensional density of $(Y - \hat{\mu}(X))^2 /
   \hat{\sigma}^2(X)$, using kernel smoothing to obtain $\hat{\rho}(Y)$.
4. Construct the estimated conditional density $p_n(Y \mid X) = \hat{\rho}((Y
   - \hat{\mu}(X)) / \hat{\sigma}(X))$.

This algorithm sketch encompasses two forms of this CDE approach, which diverge
at the second step above. To simplify the approach, one may elect to estimate
only the conditional mean $\mu(X)$ via a regression procedure, leaving the
conditional variance to be assumed constant constant (i.e., estimated simply as
the marginal mean of $\E[(Y - \hat{\mu}(X))^2]$). This class of CDE approaches
have _homoscedastic error_ based on the variance assumption made. When
additional flexibility is required, one may also estimate the conditional
variance $\sigma^2(X)$ via the residuals of the estimated conditional mean, that
is, estimating instead the conditional mean $\E[(Y - \hat{\mu}(X))^2 \mid X]$,
where the candidate algorithm $f_{\sigma}$ is used to evaluate the expectation.
These two approaches have been implemented in `sl3`, in the learner
`Lrnr_density_semiparametric`, for which the `mean_learner` argument specifies
$f_{\mu}$ and the (optional) `var_learner` argument specific $f_{\sigma}$. We
demonstrate CDE with this approach below.

```{r cde_using_locscale, eval = FALSE}
# TODO: fix code block and flesh out into demo

# semiparametric density estimator with homoscedastic errors (HOSE)
hose_hal_lrnr <- make_learner(Lrnr_density_semiparametric,
  mean_learner = hal_lrnr
)
# semiparametric density estimator with heteroscedastic errors (HESE)
hese_rf_glm_lrnr <- make_learner(Lrnr_density_semiparametric,
  mean_learner = rf_lrnr,
  var_learner = fglm_lrnr
)

# SL for the conditional treatment density
sl_dens_lrnr <- Lrnr_sl$new(
  learners = list(hose_hal_lrnr, hese_rf_glm_lrnr),
  metalearner = Lrnr_solnp_density$new()
)
```

## Conditional Density Estimation via Pooled Hazard Regression

When there is very limited information about the form of the conditional density
function itself, making it hard to justify assumptions it belonging to a
location-scale family, an alternative family of CDE approach can be leveraged.
In this vein, @diaz2011super proposed constructing conditional density
estimators (specifically in the context of estimating the generalized propensity
score) by a combination of pooled regression and exploitation of the
relationship between the (conditional) hazard and density functions. To develop
their CDE framework, @diaz2011super proposed discretizing a given $Y \in
\mathcal{Y}$ based on a number of bins $T$ and a binning procedure (e.g.,
cutting the support set $\mathcal{Y}$ into $T$ bins of exactly the same length).
This tuning parameter $T$ conceptually corresponds to the choice of bandwidth in
classical kernel density estimation. Following this discretization step, their
proposal represents each unit by a collection of records, with the number of
records for a given $O_i$ depending on the rank of the bin (along the
discretized support) into which the corresponding $Y_i$ falls.

To take an example, an instantiation of this procedure would divide the observed
support of $Y$ into, say, $T = 4$, bins of equal length, with such a
partitioning requiring $T + 1$ cutpoints along this support to yield $T$ bins:
$[\alpha_1, \alpha_2), [\alpha_2, \alpha_3), [\alpha_3, \alpha_4), [\alpha_4,
\alpha_5]$ (n.b., the rightmost interval is fully closed while the others are
only partially closed). Next, an artificial, repeated measures dataset would be
created in which each unit $O_i$ would be represented by up to $T$ records. To
better see this structure, consider an individual unit $O_i = (Y_i, X_i)$ for
which the value $Y_i$ falls in the third bin of the four into which the support
has been partitioned (i.e., $[\alpha_3, \alpha_4)$); this unit would be
represented by three distinct records: $\{Y_{ij}, X_{ij}\}_{j=1}^3$, where
$\{\{A_{ij} = 0\}_{j=1}^2$, $A_{i3} = 1\}$ and three exact copies of $X_i$,
$\{X_{ij}\}_{j=1}^3$. This representation in terms of multiple records allows
for the conditional hazard probability of $Y_i$ falling in a given bin along the
discretized support to be evaluated via standard binary regression techniques.

In fact, this proposal reformulates the binary regression problem into a
corresponding set of hazard regressions: $\mathbb{P} (Y \in [\alpha_{t-1},
\alpha_t) \mid X) = \mathbb{P} (Y \in [\alpha_{t-1}, \alpha_t) \mid Y \geq
\alpha_{t-1}, X) \times  \prod_{j = 1}^{t -1} \{1 - \mathbb{P} (Y \in
[\alpha_{j-1}, \alpha_j) \mid Y \geq \alpha_{j-1}, X) \}$. Here, the probability
of $Y \in \mathcal{Y}$ falling in bin $[\alpha_{t-1}, \alpha_t)$ may be directly
estimated via a binary regression procedure, by re-expressing the corresponding
likelihood in terms of the likelihood of a binary variable in a dataset with
this repeated measures structure. Finally, the hazard estimates can be mapped to
density estimates by re-scaling the hazard estimates by the bin sizes $\lvert
\alpha_t - \alpha_{t-1} \rvert$, that is, $p_{n, \alpha}(Y \mid X) =
\mathbb{P}(Y \in [\alpha_{t-1}, \alpha_t) \mid X) / \lvert \alpha_t -
\alpha_{t-1} \rvert$, for $\alpha_{t-1} \leq a < \alpha_t$. We provide an
informal sketch of this algorithm below.

1. Apply a procedure to divide the observed support of $Y$, $\max(Y) - \min(Y)$,
   into $T$ bins: $[\alpha_1, \alpha_2), \ldots, [\alpha_{T-1}, \alpha_T),
   [\alpha_T, \alpha_{T+1}]$.
2. Expand the observed data into a repeated measures data structure, expressing
   each individual observation as a set of up to $T$ records, recording the
   observation ID alongside each such record. For a single unit $i$, the set of
   records takes the form $\{Y_{ij}, X_{ij}\}_{j=1}^{T_i}$, where $X_{ij}$ are
   constant in the index set $\mathcal{J}$, $Y_{ij}$ is a binary counting
   process that jumps from $0$ to $1$ at its final index (at the bin into which
   $Y_i$ falls), and $T_i \leq T$ indicates the bin along its support into which
   $Y_i$ falls.
3. Estimate the hazard probability, conditional on $X$, of bin membership
   $\mathbb{P}(Y_i \in [\alpha_{t-1}, \alpha_t) \mid X)$ using any binary
   regression estimator or appropriate machine learning algorithm. So long as
   an appropriate loss function (see above) is used, cross-validation may be
   incorporated to select any tuning parameters.
4. Rescale the conditional hazard probability estimates to the conditional
   density scale by dividing the cumulative hazard by the width of the bin into
   which $X_i$ falls, for each observation $i = 1, \ldots, n$. If the support
   set is partitioned into an equal number of bins, this amounts to rescaling
   by a constant.

A key element of this proposal is the flexibility to use any binary regression
procedure or appropriate machine learning algorithm to estimate $\prob(Y \in
[\alpha_{t-1}, \alpha_t) \mid X)$, facilitating the incorporation of flexible
regression techniques like ensemble modeling [@breiman1996stacked;
@vdl2007super]. This extreme degree of flexibility integrates perfectly with the
underlying design principles of `sl3`; however, we have not yet implemented this
approach in its full generality. For now, this approach may be used through the
[`haldensify` package](https://github.com/nhejazi/haldensify)
[@hejazi2020haldensify], which limits the original proposal by replacing the use
of arbitrary binary regression with the HAL regression function, implemented in
the [`hal9001` package](https://github.com/tlverse/hal9001) [@coyle2020hal9001;
@hejazi2020hal9001]. This CDE algorithm can be used through `sl3` via the
learner `Lrnr_haldensify`, as we demonstrate below.

```{r cde_using_pooledhaz, eval = FALSE}
# TODO: fix code block and flesh out into demo

# learners used for conditional densities for (g_n)
haldensify_lrnr <- Lrnr_haldensify$new(
  n_bins = c(5, 10),
  lambda_seq = exp(seq(-1, -10, length = 200))
)
```

## Time series forecasting
