---
output:
  pdf_document: default
  html_document: default
---

# Super Learning {#sl3-advanced}

Based on the [`sl3` `R` package](https://github.com/tlverse/sl3) by _Jeremy
Coyle, Nima Hejazi, Ivana Malenica, Rachael Phillips, and Oleg Sofrygin_.

Updated: `r Sys.Date()`

## Learning Objectives{-}
1. Estimate the cross-validated predictive performance of the ensemble super 
   learner.
2. Evaluate `sl3` variable importance metrics.
3. 
4. 

## Cross-validated super learner

We can cross-validate the SL to see how well the SL performs on unseen data, and
obtain an estimate of the cross-validated risk of the SL.

This estimation procedure requires an outer/external layer of
cross-validation, also called nested cross-validation, which involves setting
aside a separate holdout sample that we don’t use to fit the SL. This external
cross-validation procedure may also incorporate 10 folds, which is the default
in `sl3`. However, we will incorporate 2 outer/external folds of
cross-validation for computational efficiency.

We also need to specify a loss function to evaluate SL. Documentation for the
available loss functions can be found in the [`sl3` Loss Function
Reference](https://tlverse.org/sl3/reference/loss_functions.html).

```{r CVsl}
washb_task_new <- make_sl3_Task(
  data = washb_data,
  covariates = covars,
  outcome = outcome,
  folds = origami::make_folds(washb_data, fold_fun = folds_vfold, V = 2)
)
CVsl <- CV_lrnr_sl(
  lrnr_sl = sl_fit, task = washb_task_new, loss_fun = loss_squared_error
)
CVsl %>%
  kable(digits = 4) %>%
  kableExtra:::kable_styling(fixed_thead = T) %>%
  scroll_box(width = "100%", height = "300px")
```
<!-- Explain summary!!!! -->

## Variable importance measures

Variable importance can be interesting and informative. It can also be
contradictory and confusing. Nevertheless, we like it, and so do our
collaborators, so we created a variable importance function in `sl3`! The `sl3`
`importance` function returns a table with variables listed in decreasing order
of importance (i.e., most important on the first row).

The measure of importance in `sl3` is based on a risk ratio, or risk difference,
between the learner fit with a removed, or permuted, covariate and the learner
fit with the true covariate, across all covariates. In this manner, the larger
the risk difference, the more important the variable is in the prediction.

The intuition of this measure is that it calculates the risk (in terms of the
average loss in predictive accuracy) of losing one covariate, while keeping
everything else fixed, and compares it to the risk if the covariate was not
lost. If this risk ratio is one, or risk difference is zero, then losing that
covariate had no impact, and is thus not important by this measure. We do this
across all of the covariates. As stated above, we can remove the covariate and
refit the SL without it, or we just permute the covariate (faster)
and hope for the shuffling to distort any meaningful information that was
present in the covariate. This idea of permuting instead of removing saves a lot
of time, and is also incorporated in the `randomForest` variable importance
measures. However, the permutation approach is risky, so the importance function
default is to remove and refit.

Let's explore the `sl3` variable importance measurements for the `washb` data.

```{r varimp, eval = FALSE}
washb_varimp <- importance(sl_fit, loss = loss_squared_error, type = "permute")
washb_varimp %>%
  kable(digits = 4) %>%
  kableExtra:::kable_styling(fixed_thead = TRUE) %>%
  scroll_box(width = "100%", height = "300px")
```

```{r varimp-plot, out.width = "100%", eval = FALSE}
# plot variable importance
importance_plot(
  washb_varimp,
  main = "sl3 Variable Importance for WASH Benefits Example Data"
)
```
<!-- Explain summary!!!! -->

## Categorical outcome prediction

## Multivariate outcome prediction

## Super learning of a conditional density
<!--
OPTIONAL NOTES TO CONSIDER ON THIS TOPIC FROM MARK'S VIDEO LECTURE:
Suppose we want to construct a Super Learner of the conditional probability
distribution $g_0(a\mid W)=P_0(A=a\mid W)$, where $a\in {\cal A}$.
Let's denote the values of $a$ with $\{0,1,\ldots,K\}$. A valid loss function
for the conditional density is
\[
L(g)(O)=-\log g(A\mid W).\]
That is, $g_0=\arg\min_g P_0L(g)$, i.e., $g_0$ is the minimizer of the
expectation of the log-likelihood loss.

**Candidate estimators**

1. Candidate estimators based on multinomial logistic regression: To start
with, one can use existing parametric model based MLE and machine learning
algorithms in `R` that fit a multinomial regression. For example, parametric
model multinomial logistic regression is available in `R` so that one can
already build a rich library of such estimators based on  different candidate
parametric models. In addition, `polyclass()` is a multinomial logistic
regression machine learning algorithm in `R`.

2. Candidate estimators based on machine learning for multinomial logistic
regression: Secondly, one can use a machine learning algorithm such as
`polyclass()` in `R` that data adaptively fits a multinomial logistic
regression, which itself has tuning parameters, again generating a class of
candidate estimators.

3. Incorporating screening: Note that one can also marry any of these choices
with a screening algorithm, thereby creating more candidate estimators of
interest. The screening can be particularly important when there are many
variables.

4. Candidate estimators by fitting separate logistic regressions and using
post-normalization

* Code $A$ in terms of Bernoullis $B_k=I(A=k)$, $k=0,\ldots,K$.
* Construct an estimator $\bar{g}_{nk}$ of $\bar{g}_{0k}(W)\equiv P_0(B_k=1\mid
  W)$ using any of the logistic regression algorithms, for all $k=0,\ldots,K$.
* This implies an estimator
\[
g_n(a\mid W)=\frac{\bar{g}_{na}(W)}{\sum_{k=0}^K \bar{g}_{nk}(W)}.\]
* In other words, we simply normalize these separate logistic regression
estimators so that we obtain a valid conditional distribution.
* This generates an enormous amount of interesting algorithms, since we have
available the whole machine learning literature for binary outcome regression.

5. Candidate estimators by estimating the conditional "hazard" with pooled
logistic regression.
Note that
\[
g_0(a\mid W)=\lambda_0(a\mid W) S_0(a\mid W),\]
where \[
\lambda_0(a\mid W)=P_0(A=a\mid A\geq a,W),\]

and $S_0(a\mid W)=\prod_{s\leq a}(1-\lambda_0(s\mid W))$ is the conditional
survival function $P_0(A>a\mid W)$. So we have now parameterized the
conditional distribution of $A$, given $W$, by a conditional hazard
$\lambda_0(a\mid W)$: $g_0=g_{\lambda_0}$.

* We could now focus on constructing candidate estimators of
$\lambda_0(a\mid W)$, which implies candidate estimators of $g_0$.

* For every observation $A_i$, we can create $A_i+1$ rows of data
$(W,s,I(A_i=s))$, $s=0,\ldots,A_i$, $i=1,\ldots,n$. We now run a logistic
regression estimator based on the pooled data set, ignoring ID, where we
regress the binary outcome $I(A_i=s)$ on the covariates $(W,s)$.

* If one assumes a parametric model, then this is nothing else then using the
maximum likelihood estimator, demonstrating that ignoring the ID is not
inefficient.

* This defines now an estimator of $\lambda_0(s\mid W)=P_0(A=s\mid W,A\geq s)$
as a function of $(s,W)$.

* Different choices of logistic regression based estimators will define
different estimators.

* The pooling across $s$ is not very sensible if $A$ is not an ordered variable
If $A$ is categorical, we recommend to compute  a separate logistic regression
estimator of $\lambda_0(a\mid W)$ for each $a$ (i.e., stratify by $s$ in the
  above pooled data set).

* For non-categorical $A$, one could include both stratified (by level) as well
as pooled (across levels) based logistic regression estimators.

-->

## Time series forecasting
