# Causal Mediation Analysis

_Nima Hejazi_

Based on the [`tmle3mediate` `R`
package](https://github.com/tlverse/tmle3mediate) by _Nima Hejazi, James
Duncan, and David McCoy_.

Updated: `r Sys.Date()`

## Learning Objectives

1. TODO
2. TODO
3. TODO
4. TODO
5. TODO
6. TODO
7. TODO

## Introduction to Causal Mediation Analysis

A treatment often affects an outcome indirectly, through a particular pathway,
by its effect on _intermediate variables_ (mediators). Causal mediation analysis
concerns the construction and evaluation of these _indirect effects_ and their
complementary _direct effects_. Generally, the indirect effect (IE) of a
treatment on an outcome is the portion of the total effect that is found to work
_through_ mediator variables, while the direct effect often encompasses all
other components of the total effect, including both the effect of the treatment
on the outcome _and_ the effect through all paths not explicitly involving the
mediators). Identifying and quantifying the mechanisms underlying causal effects
is an increasingly desirable endeavor in public health, medicine, and the social
sciences, as such mechanistic knowledge improves understanding of both _why_ and
_how_ treatments may be effective.

While the study of mediation analysis may be traced back quite far, the field
only came into its modern form with the identification and careful study of the
natural direct and indirect effects [@robins1992identifiability;
pearl2001direct]. The natural direct effect (NDE) and the natural indirect
effect (NIE) are based on a decomposition of the average treatment effect (ATE)
in the presence of mediators [@vanderweele2015explanation]; requisite
theory for the construction of efficient estimators of these quantities only
receiving attention relatively recently [@tchetgen2012semiparametric].

## Data Structure and Notation

Consider $n$ observed units $O_1, \ldots, O_n$, where each observed data random
variable takes the form $O = (W, A, Z, Y)$, for a vector of observed covariates
$W$, a binary or continuous treatment $A$, possibly multivariate mediators $Z$,
and a binary or continuous outcome $Y$. To avoid undue assumptions, we assume
only that $O \sim \mathcal{P} \in \M$ where $\M$ is the nonparametric
statistical model defined as all continuous densities on $O$ with respect to an
arbitrary dominating measure.

We formalize the definition of our counterfactual variables using the following
non-parametric structural equation model (NPSEM):
\begin{align}
  W &= f_W(U_W) \\
  A &= f_A(W, U_A) \\
  Z &= f_Z(W, A, U_M) \\
  Y &= f_Y(W, A, Z, U_Y).
  (\#eq:npsem-mediate)
\end{align}
This set of equations
represents a mechanistic model generating the observed data $O$; furthermore,
the NPSEM encodes several fundamental assumptions. Firstly, there is an implicit
temporal ordering: $W$ occurs first, depending only on exogenous factors $U_W$;
$A$ happens next, based on both $W$ and exogenous factors $U_A$; then come the
mediators $Z$, which depend on $A$, $W$, and another set of exogenous factors
$U_Z$; and finally appears the outcome $Y$. We assume neither access to the set
of exogenous factors $\{U_W, U_A, U_Z, U_Y\}$ nor knowledge of the forms of the
deterministic generating functions $\{f_W, f_A, f_Z, f_Y\}$. The NPSEM
corresponds to the following DAG:

```{r mediation-DAG}
library(dagitty)
library(ggdag)

# make DAG by specifying dependence structure
dag <- dagitty(
  "dag {
    W -> A
    W -> Z
    W -> Y
    A -> Z
    A -> Y
    Z -> Y
    W -> A -> Y
    W -> A -> Z -> Y
  }"
)
exposures(dag) <- c("A")
outcomes(dag) <- c("Y")
tidy_dag <- tidy_dagitty(dag)

# visualize DAG
ggdag(tidy_dag) +
  theme_dag()
```

The likelihood of the data $O$ admits a factorization, wherein, for $p_0^O$,
the density of $O$ with respect to the product measure, the density evaluated
on a particular observation $o$ may be a written
\begin{equation}
  p_0^O(x) = q^O_{0,Y}(y \mid Z = z A = a, W = w)
    q^O_{0,Z}(z \mid A = a, W = w)
    q^O_{0,A}(a \mid W = w)
    q^O_{0,W}(w),
  (\#eq:likelihood-factorization-mediate)
\end{equation}
where $q_{0, Y}$ is the conditional density of $Y$ given $(Z, A, W)$, $q_{0, Z}$
is the conditional density of $Z$ given $(A, W)$, $q_{0, A}$ is the conditional
density of $A$ given $W$, and $q_{0, W}$ is the density of $W$. Further, for
ease of notation, we let $\bar{Q}_Y(Z, A, W) = \E[Y \mid Z, A, W]$,
$\bar{Q}_Z(A, W) = \P[Z \mid A, W]$ (later re-parametrized to $e(A \mid Z, W) =
\P(A \mid Z, W)$), $g(A \mid W) = \P(A \mid W)$, and $q_W$ the marginal
distribution of $W$.

Finally, note that we have explicitly excluded potential confounders of the
mediator-outcome relationship affected by exposure (i.e., variables affected by
$A$ and affecting both $Z$ and $Y$). Mediation analysis in the presence of such
variables is exceptionally challenging [@avin2005identifiability]; thus, most
efforts to develop definitions of causal direct and indirect effects explicitly
disallowed such a form of confounding. While we will not discuss the matter
here, the interested reader may consult recent advances in the vast literature
on causal mediation analysis, among them @diaz2020nonparametric and
@hejazi2021nonparametric.

## The Natural Direct and Indirect Effects

### Decomposing the Average Treatment Effect

The natural direct and indirect effects arise from a decomposition of the ATE:
\begin{equation*}
  \E[Y(1) - Y(0)] =
    \underbrace{\E[Y(1, Z(0)) - Y(0, Z(0))]}_{NDE} +
    \underbrace{\E[Y(1, Z(1)) - Y(1, Z(0))]}_{NIE}.
\end{equation*}
In particular, the natural indirect effect (NIE) measures the effect of the
treatment $A \in \{0, 1\}$ on the outcome $Y$ through the mediators $Z$, while
the natural direct effect (NDE) measures the effect of the treatment on the
outcome _through all other paths_.

### The Natural Direct Effect

The NDE is defined as
\begin{equation*}
  \Psi_{NDE} = \E[Y(1, Z(0)) - Y(0, Z(0))]
  \overset{\text{rand.}}{=} \sum_w \sum_z
  [\underbrace{\E(Y \mid A = 1, z, w)}_{\bar{Q}_Y(A = 1, z, w)} -
  \underbrace{\E(Y \mid A = 0, z, w)}_{\bar{Q}_Y(A = 0, z, w)}] \times
  \underbrace{p(z \mid A = 0, w)}_{Q_Z(0, w))} \underbrace{p(w)}_{q_W},
\end{equation*}
where the likelihood factors $p(z \mid A = 0, w)$ and $p(w)$ (among other
conditional densities) arise from a factorization of the joint likelihood:
\begin{equation*}
  p(w, a, z, y) = \underbrace{p(y \mid w, a, z)}_{Q_Y(A, W, Z)}
  \underbrace{p(z \mid w, a)}_{Q_Z(Z \mid A, W)}
  \underbrace{p(a \mid w)}_{g(A \mid W)}
  \underbrace{p(w)}_{Q_W}.
\end{equation*}

The process of estimating the NDE begins by constructing $\bar{Q}_{Y, n}$, an
estimate of the outcome mechanism $\bar{Q}_Y(Z, A, W) = \E \{Y \mid Z, A,
W\}$ (i.e., the conditional mean of $Y$, given $Z$, $A$, and $W$). With an
estimate of this conditional expectation in hand, predictions of the
counterfactual quantities $\bar{Q}_Y(Z, 1, W)$ (setting $A = 1$) and, likewise,
$\bar{Q}_Y(Z, 0, W)$ (setting $A = 0$) can readily be obtained. We denote the
difference of these counterfactual quantities $\bar{Q}_{\text{diff}}$, i.e.,
$\bar{Q}_{\text{diff}} = \bar{Q}_Y(Z, 1, W) - \bar{Q}_Y(Z, 0, W)$.
$\bar{Q}_{\text{diff}}$ represents the difference in the conditional mean of $Y$
attributable to changes in $A$ while keeping $Z$ and $W$ at their _natural_
(i.e., observed) values.

The estimation procedure treats $\bar{Q}_{\text{diff}}$ itself as a nuisance
parameter, regressing its estimate $\bar{Q}_{\text{diff}, n}$ on $W$, among
control observations only (i.e., those for whom $A = 0$ is observed); the goal
of this step is to remove part of the marginal impact of $Z$ on
$\bar{Q}_{\text{diff}}$, since $W$ is a parent of $Z$. Regressing this
difference on $W$ among the controls recovers the expected
$\bar{Q}_{\text{diff}}$, had all individuals been set to the control condition
$A = 0$. Any residual additive effect of $Z$ on $\bar{Q}_{\text{diff}}$ is
removed during the TML estimation step using the auxiliary (or "clever")
covariate, which accounts for the mediators $Z$. This auxiliary covariate takes
the form
\begin{equation*}
  C_Y(Q_Z, g)(O) = \Bigg\{\frac{\mathbb{I}(A = 1)}{g(1 \mid W)}
  \frac{Q_Z(Z \mid 0, W)}{Q_Z(Z \mid 1, W)} -
  \frac{\mathbb{I}(A = 0)}{g(0 \mid W)} \Bigg\}.
\end{equation*}
Breaking this down, $\frac{\mathbb{I}(A = 1)}{g(1 \mid W)}$ is the inverse
propensity score weight for $A = 1$ and, likewise, $\frac{\mathbb{I}(A = 0)}
{g(0 \mid W)}$ is the inverse propensity score weight for $A = 0$. The middle
term is the ratio of the conditional densities of the mediator under the control
($A = 0$) and treatment ($A = 1$) conditions.

This subtle appearance of a ratio of conditional densities is concerning --
unfortunately, tools to estimate such quantities are sparse in the statistics
literature, and the problem is still more complicated (and computationally
taxing) when $Z$ is high-dimensional. As only the ratio of these conditional
densities is required, a convenient re-parametrization may be achieved, that is,
\begin{equation*}
  \frac{p(A = 0 \mid Z, W) g(0 \mid W)}{p(A = 1 \mid Z, W) g(1 \mid W)}.
\end{equation*}
Going forward, we will denote this re-parameterized conditional probability
$e(A \mid Z, W) := p(A \mid Z, W)$. Similar re-parameterizations have been used
in @zheng2012targeted and @tchetgen2013inverse. This is particularly useful
since this reformulation reduces the problem to one concerning only the
estimation of conditional means, opening the door to the use of a wide range of
machine learning algorithms (e.g., most of those in
[`sl3`](https://github.com/tlverse/sl3)).

Underneath the hood, the counterfactual outcome difference
$\bar{Q}_{\text{diff}}$ and $e(A \mid Z, W)$, the conditional probability of $A$
given $Z$ and $W$, are used in constructing the auxiliary covariate for TML
estimation. These nuisance parameters play an important role in the
bias-correcting _TMLE-update step_.

### The Natural Indirect Effect

Derivation and estimation of the NIE is analogous to that of the NDE.  The NIE
is the effect of $A$ on $Y$ _only through the mediator(s) $Z$_.  This quantity
-- known as the (additive) natural indirect effect $\mathbb{E}(Y(Z(1), 1)
- \mathbb{{E}(Y(Z(0), 1)$ -- corresponds to the difference of the conditional
expectation of $Y$ given $A = 1$ and $Z(1)$ (the values the mediator would take
under $A = 1$) and the conditional expectation of $Y$ given $A = 1$ and $Z(0)$
(the values the mediator would take under $A = 0$).

As with the NDE, the re-parameterization trick can be used to estimate
$\mathbb{E}(A \mid Z, W)$, avoiding estimation of a possibly multivariate
conditional density. However, in this case, the mediated mean outcome
difference, denoted $\Psi_Z(Q)$, is instead estimated as follows
\begin{equation*}
  \Psi_{NIE}(Q) = \mathbb{E}_{QZ}(\Psi_{NIE, Z}(Q)(1, W) -
    \Psi_{NIE, Z}(Q)(0, W))
\end{equation*}

Phrased plainly, this uses $\bar{Q}_Y(Z, 1, W)$ (the predicted values for $Y$
given $Z$ and $W$ when $A = 1$) and regresses this vector on $W$ among the
treated units (for whom $A = 1$ is observed) to obtain the conditional mean
$\Psi_{NIE, Z}(Q)(1, W)$. Performing the same procedure, but now regressing
$\bar{Q}_Y(Z, 1, W)$ on $W$ among the control units (for whom $A = 0$ is
observed) yields $\Psi_{NIE,Z}(Q)(0, W)$. The difference of these two estimates
is the NIE and can be thought of as the additive marginal effect of treatment on
the conditional expectation of $Y$ given $W$, $A = 1$, $Z$ through its effects
on $Z$. So, in the case of the NIE, our estimate $\psi_n$ is slightly different,
but the same quantity $e(A \mid Z, W)$ comes into play as the auxiliary
covariate.

## The Population Intervention Direct and Indirect Effects

At times, the natural direct and indirect effects may prove too limiting, as
these effect definitions are based on _static interventions_ (i.e., setting $A =
0$ or $A = 1$), which may be unrealistic for real-world interventions. In such
cases, one may turn instead to the population intervention direct effect (PIDE)
and the population intervention indirect effect (PIIE), which are based on
decomposing the effect of the population intervention effect (PIE) of flexible
stochastic interventions [@diaz2020causal].

A particular type of stochastic intervention well-suited to working with binary
treatments is the _incremental propensity score intervention_ (IPSI), first
proposed by @kennedy2017nonparametric. Such interventions do not
deterministically set the treatment level of an observed unit to a fixed
quantity (i.e., setting $A = 1$), but instead _alter the odds of receiving the
treatment_ by a fixed amount ($0 \leq \delta \leq \infty$) for each individual.
In the case of the `mma` dataset, we will proceed by considering an IPSI that
modulates the odds of participating in a sports team by $\delta = 2$.  Such an
intervention may be interpreted (hypothetically) as the effect of a school
program that motivates children to participate in sports teams (e.g., as in an
encouragement design). To exemplify our approach, we postulate a motivational
intervention that _doubles the odds_ (i.e., $\delta = 2$) of participating in a
sports team for each individual:

```{r delta_ipsi, message=FALSE, warning=FALSE}
delta_ipsi <- 2
```

### Decomposing the Population Intervention Effect

We may decompose the population intervention effect (PIE) in terms of the
_population intervention direct effect_ (PIDE) and the _population
intervention indirect effect_ (PIIE):
\begin{equation*}
  \mathbb{E}\{Y(A_\delta)\} - \mathbb{E}Y =
    \overbrace{\mathbb{E}\{Y(A_\delta, Z(A_\delta))
      - Y(A_\delta, Z)\}}^{\text{PIIE}} +
    \overbrace{\mathbb{E}\{Y(A_\delta, Z) - Y(A, Z)\}}^{\text{PIDE}}.
\end{equation*}

This decomposition of the PIE as the sum of the population intervention direct
and indirect effects has an interpretation analogous to the corresponding
standard decomposition of the average treatment effect. In the sequel, we will
compute each of the components of the direct and indirect effects above using
appropriate estimators as follows

* For $\mathbb{E}\{Y(A, Z)\}$, the sample mean $\frac{1}{n}\sum_{i=1}^n Y_i$ is
  consistent;
* for $\mathbb{E}\{Y(A_{\delta}, Z)\}$, a TML estimator for the effect of a
  joint intervention altering the treatment mechanism but not the mediation
  mechanism, based on the proposal in @diaz2020causal; and,
* for $\mathbb{E}\{Y(A_{\delta}, Z_{A_{\delta}})\}$, an efficient estimator for
  the effect of a joint intervention altering both the treatment and mediation
  mechanisms, as proposed in @kennedy2017nonparametric and implemented in the
  [`npcausal` R package](https://github.com/ehkennedy/npcausal).

### Estimating the Effect Decomposition Term

As described by @diaz2020causal, the statistical functional identifying the
decomposition term that appears in both the PIDE and PIIE
$\mathbb{E}\{Y(A_{\delta}, Z)\}$, which corresponds to altering the treatment
mechanism while keeping the mediation mechanism fixed, is
\begin{equation*}
  \theta_0(\delta) = \int m_0(a, z, w) g_{0,\delta}(a \mid w) p_0(z, w)
    d\nu(a, z, w),
\end{equation*}
for which a TML estimator is available. The corresponding _efficient influence
function_ (EIF) with respect to the nonparametric model $\mathcal{M}$ is
$D_{\eta,\delta}(o) = D^Y_{\eta,\delta}(o)
+ D^A_{\eta,\delta}(o) + D^{Z,W}_{\eta,\delta}(o) - \theta(\delta)$.

The TML estimator may be computed basd on the EIF estimating equation and may
incorporate cross-validation [@zheng2011cross; @chernozhukov2018double] to
circumvent possibly restrictive entropy conditions (e.g., Donsker class). The
resultant estimator is
\begin{equation*}
  \hat{\theta}(\delta) = \frac{1}{n} \sum_{i = 1}^n D_{\hat{\eta}_{j(i)},
  \delta}(O_i) = \frac{1}{n} \sum_{i = 1}^n \left\{ D^Y_{\hat{\eta}_{j(i)},
  \delta}(O_i) + D^A_{\hat{\eta}_{j(i)}, \delta}(O_i) +
  D^{Z,W}_{\hat{\eta}_{j(i)}, \delta}(O_i) \right\},
\end{equation*}
which is implemented in `tmle3mediate` (a one-step estimator is also avaialble,
in the [`medshift` R package](https://github.com/nhejazi/medshift)). We
demonstrate the use of `tmle3mediate` to obtain $\mathbb{E}\{Y(A_{\delta}, Z)\}$
via its TML estimator:



















```{r load_data, message=FALSE, warning=FALSE}
# preliminaries
library(dplyr)
library(tidyr)
library(sl3)
library(tmle3)
library(tmle3mediate)

# set a seed
set.seed(429153)
```

For the analysis of this observational dataset, we focus on the effect of
participating in a sports team (`sports`) on the BMI of children (`bmi`), taking
several related covariates as mediators (`snack`, `exercises`, `overweigh`) and
all other collected covariates as potential confounders. Considering an NPSEM,
we separate the observed variables from the dataset into their corresponding
nodes as follows:

```{r npsem, message=FALSE, warning=FALSE}
exposure <- "sports"
outcome <- "bmi"
mediators <- c("snack", "exercises", "overweigh")
covars <- setdiff(colnames(weight_behavior_complete),
                  c(exposure, outcome, mediators))
node_list <- list(
  W = covars,
  A = exposure,
  Z = mediators,
  Y = outcome
)
```

Here the `node_list` encodes the parents of each node -- for example, $Z$ (the
mediators) have parents $A$ (the treatment) and $W$ (the baseline confounders),
and $Y$ (the outcome) has parents $Z$, $A$, and $W$.

To use the natural (in)direct effects for mediation analysis, we decompose the
ATE into its corresponding direct and indirect effect components. Throughout,
we'll make use of the (semiparametric efficient) TML estimators of
@zheng2012targeted, which allow for flexible ensemble machine learning to be
incorporated into the estimation of nuisance parameters. For this, we rely on
the [`sl3` R package](https://tlverse.org/sl3) [@coyle2020sl3], which provides
an implementation of machine learning pipelines and the Super Learner algorithm;
for more on the `sl3` R package, consider consulting [this chapter of the
`tlverse` handbook](https://tlverse.org/tlverse-handbook/sl3.html). Below, we
construct an ensemble learner using a handful of popular machine learning
algorithms:

```{r make_sl_learners, message=FALSE, warning=FALSE}
# SL learners used for continuous data (the nuisance parameter M)
enet_contin_learner <- Lrnr_glmnet$new(
  alpha = 0.5, family = "gaussian", nfolds = 3
)
lasso_contin_learner <- Lrnr_glmnet$new(
  alpha = 1, family = "gaussian", nfolds = 3
)
fglm_contin_learner <- Lrnr_glm_fast$new(family = gaussian())
mean_learner <- Lrnr_mean$new()
contin_learner_lib <- Stack$new(
  enet_contin_learner, lasso_contin_learner, fglm_contin_learner, mean_learner
)
sl_contin_learner <- Lrnr_sl$new(learners = contin_learner_lib,
                                 metalearner = Lrnr_nnls$new())

# SL learners used for binary data (nuisance parameters G and E in this case)
enet_binary_learner <- Lrnr_glmnet$new(
  alpha = 0.5, family = "binomial", nfolds = 3
)
lasso_binary_learner <- Lrnr_glmnet$new(
  alpha = 1, family = "binomial", nfolds = 3
)
fglm_binary_learner <- Lrnr_glm_fast$new(family = binomial())
binary_learner_lib <- Stack$new(
  enet_binary_learner, lasso_binary_learner, fglm_binary_learner, mean_learner
)
logistic_metalearner <- make_learner(Lrnr_solnp,
                                     metalearner_logistic_binomial,
                                     loss_loglik_binomial)
sl_binary_learner <- Lrnr_sl$new(learners = binary_learner_lib,
                                 metalearner = logistic_metalearner)

# create list for treatment and outcome mechanism regressions
learner_list <- list(
  Y = sl_contin_learner,
  A = sl_binary_learner
)
```

## Estimating the Natural Indirect Effect

We demonstrate calculation of the NIE using `tmle3mediate` below, starting by
instantiating a "Spec" object that encodes exactly which learners to use for the
nuisance parameters $e(A \mid Z, W)$ and $\Psi_Z$. We then pass our Spec object
to the `tmle3` function, alongside the data, the node list (created above), and
a learner list indicating which machine learning algorithms to use for
estimating the nuisance parameters based on $A$ and $Y$.

```{r NIE, message=FALSE, warning=FALSE}
tmle_spec_NIE <- tmle_NIE(
  e_learners = Lrnr_cv$new(lasso_binary_learner, full_fit = TRUE),
  psi_Z_learners = Lrnr_cv$new(lasso_contin_learner, full_fit = TRUE),
  max_iter = 1
)
weight_behavior_NIE <- tmle3(
  tmle_spec_NIE, weight_behavior_complete, node_list, learner_list
)
weight_behavior_NIE
```

## Estimating the Natural Direct Effect

An analogous procedure applies for estimation of the NDE, only replacing the
Spec object for the NIE with `tmle_spec_NDE` to define learners for the NDE
nuisance parameters:

```{r NDE, message=FALSE, warning=FALSE}
tmle_spec_NDE <- tmle_NDE(
  e_learners = Lrnr_cv$new(lasso_binary_learner, full_fit = TRUE),
  psi_Z_learners = Lrnr_cv$new(lasso_contin_learner, full_fit = TRUE),
  max_iter = 1
)
weight_behavior_NDE <- tmle3(
  tmle_spec_NDE, weight_behavior_complete, node_list, learner_list
)
weight_behavior_NDE
```

### PIDE PIIE

TODO

```{r efficient_est, message=FALSE, warning=FALSE}
# instantiate tmle3 spec for stochastic mediation
tmle_spec_pie_decomp <- tmle_medshift(
  delta = delta_ipsi,
  e_learners = Lrnr_cv$new(lasso_binary_learner, full_fit = TRUE),
  phi_learners = Lrnr_cv$new(lasso_contin_learner, full_fit = TRUE)
)

# compute the TML estimate
weight_behavior_pie_decomp <- tmle3(
  tmle_spec_pie_decomp, weight_behavior_complete, node_list, learner_list
)
weight_behavior_pie_decomp
```

## Estimating the Population Intervention Direct Effect

Recall that, based on the decomposition outlined previously, the population
intervention direct effect may be denoted $\beta_{\text{PIDE}}(\delta) =
\theta_0(\delta) - \mathbb{E}Y$. Thus, an estimator of the PIDE,
$\hat{\beta}_{\text{PIDE}}(\delta)$ may be expressed as a composition of
estimators of its constituent parameters:
\begin{equation*}
  \hat{\beta}_{\text{PIDE}}({\delta}) = \hat{\theta}(\delta) -
  \frac{1}{n} \sum_{i = 1}^n Y_i.
\end{equation*}

Based on the above, we may construct an estimator of the PIDE using quantities
already computed. To do this, we need only apply the delta method, available
from the [`tmle3` package](https://github.com/tlverse/tmle3).

<!--
```{r pide_delta, message=FALSE, warning=FALSE, eval=FALSE}
tmle_task <- tmle_spec_pie_decomp$make_tmle_task(
  weight_behavior_complete, node_list
)
initial_likelihood <- tmle_spec_pie_decomp$make_initial_likelihood(
  tmle_task, learner_list
)
```
-->

## Exercises

### Review of Key Concepts

1. Additivity of the natural (in)direct effects and the ATE.

2. Working with continuous mediators.

3. TODO

### The Ideas in Action

1. TODO

2. TODO

3. TODO

### Advanced Topics

1. TODO

2. TODO

3. TODO

<!--
## Appendix

### Exercise solutions
-->
